---
title: "Project 1"
---

```{r render, echo = FALSE, eval = FALSE}
rmarkdown::render("README.Rmd",
                  output_format = "github_document",
                  output_options = list(
                      toc = TRUE,
                      html_preview=FALSE) 
)
```

# Lets Get Started: OMDb API Key

In order to access the OMDb API, you need to get a free [api key](http://www.omdbapi.com/apikey.aspx). In the rest of this document, "mykey" refers to your OMDb API key.

You should also "turn on" these packages by running the code below. If you don't have them installed yet, run `install.packages()` with the package in quotes. For example, to install `tidyverse`, you would run `install.packages("tidyverse")`.

```{r 7_03_6_22_2022,message=FALSE,warning=FALSE}
library(httr) #this package will help use use the URL we built to get information from the OMDb API
library(jsonlite) #this package will help us convert the data we get from the OMDb API to a more usable format
library(tidyverse) #this package will help us work with our nicely formatted data.
library(lubridate) #this package will help us create dates 
```

In order to get information from the OMDb API, we have to build a URL with our search criteria. It's similar to doing a Google search. There are two ways to build a URL: "By ID or Title" or "By Search".

# Build URL for One Movie Title

Let's say you have a movie title in mind, like Star Wars (1977). Here's a function you can use to get data from the OMDb API about Star Wars:

Note: the parameter “type” has three options: movie, series, and episode. If “type” is not specified, it will give everything (including movies, series, and episodes). I’m making the default for “type” be “movie”, but you can change this if you want.

```{r search_by_title_Star_Wars}
search_by_title <- function(mykey,title,type="movie"){
  base_url <- paste0("http://www.omdbapi.com/?apikey=",mykey)
  info_url <- paste0("&t=",title,"&type=",type) 
  full_url <- paste0(base_url, info_url)
  full_url <- gsub(full_url, pattern = " ", replacement = "%20")
  
  movie_api_call <- GET(full_url)
  movie_api_call_char <- rawToChar(movie_api_call$content)
  movie_JSON <- jsonlite::fromJSON(movie_api_call_char, flatten = TRUE) 
  movie_JSON <- as.data.frame(movie_JSON)
  tibble_movie_JSON <- as_tibble(movie_JSON)
  return(tibble_movie_JSON)
}
```

You should run the function like this:

```{r 3_50_6_21_2022,eval=FALSE}
search_by_title("mykey","star_wars",type="movie")
```


You should get a tibble that looks like this:

```{r example_search_by_title_Star_Wars, eval=TRUE, echo=FALSE}
search_by_title("5c7f9206","star_wars",type="movie")
```

# Build URL for One Movie Title and One Date

If you don't specify a date, the OMDb API will give the first result. So, since Star Wars (1977) was the first Star Wars movie ever made, it gives Star Wars (1977) as the result. But, what if you wanted a different Star Wars movie like Star Wars: Episode VII - The Force Awakens (2015)? You can use this funciton:

```{r 7_11_6_22_2022}
search_by_title_and_date <- function(mykey,title,type="movie",date){
  base_url <- paste0("http://www.omdbapi.com/?apikey=",mykey)
  info_url <- paste0("&t=",title,"&type=",type,"&y=",date) 
  full_url <- paste0(base_url, info_url)
  full_url <- gsub(full_url, pattern = " ", replacement = "%20")
  
  movie_api_call <- GET(full_url)
  movie_api_call_char <- rawToChar(movie_api_call$content)
  movie_JSON <- jsonlite::fromJSON(movie_api_call_char, flatten = TRUE) 
  movie_JSON <- as.data.frame(movie_JSON)
  tibble_movie_JSON <- as_tibble(movie_JSON)
  return(tibble_movie_JSON)
}
```


You should run the function like this:

```{r 8_00_6_22_2022,eval=FALSE}
search_by_title_and_date("mykey","star_wars",type="movie",date=2015)
```

Get a tibble like this:

```{r 7_13_6_22_2022,eval=TRUE,echo=FALSE}
search_by_title_and_date("5c7f9206","star_wars",type="movie",date=2015)
```

# Build URL for One IMDb ID

Lets say you have a valid IMDb ID. You can find an IMDb ID by searching for a title on the [IMDb website](www.imdb.com). After you find a movie you like, the IMDb ID will be in the URL. For example, the URL for the IMDb page for Star Wars: Episode V - The Empire Strikes Back (1980) is https://www.imdb.com/title/tt0080684/?ref_=nv_sr_srsg_0. Therefore, its IMDb ID is tt0080684.

Here's a function you can use if you have a valid IMDb ID:

```{r search_by_IMDb_ID_Star_Wars, message=FALSE,warning=FALSE}
library(tidyverse)
library(jsonlite)
library(httr)
search_by_IMDb_ID <- function(mykey,IMDb_ID,type="movie"){
  base_url <- paste0("http://www.omdbapi.com/?apikey=",mykey)
  info_url <- paste0("&i=",IMDb_ID,"&type=",type) 
  full_url <- paste0(base_url, info_url)
  full_url <- gsub(full_url, pattern = " ", replacement = "%20")
  
  movie_api_call <- GET(full_url)
  movie_api_call_char <- rawToChar(movie_api_call$content)
  movie_JSON <- jsonlite::fromJSON(movie_api_call_char, flatten = TRUE) 
  movie_JSON <- as.data.frame(movie_JSON)
  tibble_movie_JSON <- as_tibble(movie_JSON)
  return(tibble_movie_JSON)
}
```

You should run the function like this:

```{r 3_54_6_21_2022,eval=FALSE}
search_by_IMDb_ID("mykey","tt0080684",type="movie")
```

You should get a tibble that looks like this:

```{r example_search_by_IMDb_ID_Star_Wars, eval=TRUE, echo=FALSE}
search_by_IMDb_ID("5c7f9206","tt0080684",type="movie")
```

# Build URL to Search for More than One Title in a Series

Let's say you wanted to get all of the titles for all of the Star Wars movies. You would then need to build your URL "By Search" instead. Here's a function you can use if you wanted to search for multiple movie titles:

```{r by_search, message=FALSE,warning=FALSE}
by_search_series <- function(mykey,title,type="movie"){
  base_url <- paste0("http://www.omdbapi.com/?apikey=",mykey)
  info_url <- paste0("&s=",title,"&type=",type) 
  full_url <- paste0(base_url, info_url)
  full_url <- gsub(full_url, pattern = " ", replacement = "%20")
  
  movie_api_call <- GET(full_url)
  movie_api_call_char <- rawToChar(movie_api_call$content)
  movie_JSON <- jsonlite::fromJSON(movie_api_call_char, flatten = TRUE) 
  movie_JSON <- as.data.frame(movie_JSON)
  tibble_movie_JSON <- as_tibble(movie_JSON)
  return(tibble_movie_JSON)
}
```

You should run the function like this:

```{r 3_57_6_21_2022,eval=FALSE}
by_search_series("mykey","star_wars",type="movie")
```

You should get a tibble that looks like this:

```{r example_by_search, eval=TRUE, echo=FALSE}
by_search_series("5c7f9206","star_wars",type="movie")
```

# Get the data!

That's great! Now, lets get the data for all of the titles:

```{r get_data}
mat=NULL
get_data_series <- function(mykey,title){
  temp_table <- by_search_series(mykey,title,type="movie")
  list_of_titles <- unique(temp_table$Search.Title)
  
  for(movie_title in list_of_titles){
  table <- search_by_title(mykey,movie_title,type="movie")
  mat=rbind(mat,table)
  }
  return(mat)
}
```

You should run the function like this:

```{r 8_54_6_22_2022,eval=FALSE}
get_data_series("mykey","star_wars")
```

You should get a tibble that looks like this:

```{r example_get_data,eval=TRUE,echo=FALSE}
get_data_series("5c7f9206","star_wars")
```

# Build URL to search for More than One Title or Series

Now, what if you want to get all of the data for all of the Star Wars movies and all of the Indiana Jones movies. The function below can handle a list of several titles or one title.

```{r test_3_35_6_20_2022}
mat=NULL
by_search_one_or_more_titles <- function(mykey,title,type="movie"){
 if(length(title)<=1){ 
  base_url <- paste0("http://www.omdbapi.com/?apikey=",mykey)
    info_url <- paste0("&s=",title,"&type=",type) 
    full_url <- paste0(base_url, info_url)
    full_url <- gsub(full_url, pattern = " ", replacement = "%20")
    
    movie_api_call <- GET(full_url)
    movie_api_call_char <- rawToChar(movie_api_call$content)
    movie_JSON <- jsonlite::fromJSON(movie_api_call_char, flatten = TRUE) 
    movie_JSON <- as.data.frame(movie_JSON)
    movie_JSON <- as_tibble(movie_JSON)
    return(movie_JSON)
 }
  if(length(title)>1){
  for(i in title){
  base_url <- paste0("http://www.omdbapi.com/?apikey=",mykey)
  info_url <- paste0("&s=",i,"&type=",type) 
  full_url <- paste0(base_url, info_url)
  full_url <- gsub(full_url, pattern = " ", replacement = "%20")
  
  movie_api_call <- GET(full_url)
  movie_api_call_char <- rawToChar(movie_api_call$content)
  movie_JSON <- jsonlite::fromJSON(movie_api_call_char, flatten = TRUE) 
  movie_JSON <- as.data.frame(movie_JSON)
  mat=rbind(mat,movie_JSON)
  mat=as_tibble(mat)
  }
    }
  return(mat)
}
```

You can use this function to look for multiple titles or just one. For example, if you wanted to search for both Star Wars and Indiana Jones, you would run the function like this:

```{r ex_test_3_50_6_20_2022,eval=FALSE}
by_search_one_or_more_titles("mykey",c("star_wars","indiana_jones"),type="movie")
```

You would get the tibble below: 

```{r ex_test_3_35_6_20_2022,eval=TRUE,echo=FALSE}
by_search_one_or_more_titles("5c7f9206",c("star_wars","indiana_jones"),type="movie")
```

If you wanted to search for one title, like Spider-Man, you would run the function like this:

```{r ex_test_3_54_6_20_2022,eval=FALSE}
by_search_one_or_more_titles("mykey","spider-man",type="movie")
```

You would get this tibble:

```{r 9_04_6_22_2022,eval=TRUE,echo=FALSE}
by_search_one_or_more_titles("5c7f9206","spider-man",type="movie")
```

Now, lets get all of the data for both Star Wars and Indiana Jones:

```{r 9_03_6_22_2022}
mat=NULL
get_data_one_or_more_titles <- function(mykey,title){
  temp_table <- by_search_one_or_more_titles(mykey,title,type="movie")
  list_of_titles <- unique(temp_table$Search.Title)
  
  for(movie_title in list_of_titles){
  table <- search_by_title(mykey,movie_title,type="movie")
  mat=rbind(mat,table)
  }
  return(mat)
}
```

You would run the function like this:

```{r 9_05_6_22_2022,eval=FALSE}
get_data_one_or_more_titles("mykey",c("star_wars","indiana_jones"))
```


```{r test_5_11_6_20_2022,echo=FALSE}
get_data_one_or_more_titles("5c7f9206",c("star_wars","indiana_jones"))
```

# Let's make a data set!

First, I'm going to make two lists of movies:

```{r 10_04_6_22_2022}
#for these movies, I just want the first result it gives me:
titles <- c("casablanca",
            "the_wizard_of_oz",
            "it's_a_wonderful_life",
            "goodfellas",
            "taxi_driver",
            "psycho",
            "singin_in_the_rain",
            "2001:_a_space_odyssey",
            "vertigo")

#for these movies, I want all of them in the series:
series <- c("the_godfather",
            "star_wars",
            "alien",
            "fast_and_furious",
            "final_destination",
            "friday_the_13th")
```

Here is the function I'm going to use to get all of the data for all of my movies:

```{r 11_50_6_22_2022}
mat1=NULL
mat2=NULL
mat3=NULL
get_data_titles_and_series <- function(mykey,titles,series){
  for(i in titles){
    temp_table <- search_by_title(mykey,i,type="movie")
      mat1=rbind(mat1,temp_table)
  }
  for(j in series){
    temp_table <- by_search_series(mykey,j,type="movie")
    list_of_titles <- unique(temp_table$Search.Title)
    for(movie_title in list_of_titles){
      table2 <- search_by_title(mykey,movie_title,type="movie")
      mat2=rbind(mat2,table2)
    }
  }
  mat3=rbind(mat3,mat1,mat2)
  return(mat3)
}
```

I'm going to run the function like this:

```{r 11_51_6_22_2022,eval=FALSE}
get_data_titles_and_series("mykey",titles,series)
```

Here is the tibble I get:

```{r 11_52_6_22_2022, echo=FALSE}
get_data_titles_and_series("5c7f9206",titles,series)
```

Before we make this data more usable, we need to write a few more helper functions:

```{r 7_46_6_23_2022}
#this will help us convert the Ratings.Value column to numeric
parse_number <- function(S){
  if(grepl("/", S)){
    A<-str_split(S, "/")
    A<-as.numeric(unlist(A))
    A<-A[[1]]/A[[2]]
    A<-A*100
  } else {
    A<-as.numeric(gsub("%","",S))
  }
  return(A)
}

#this will help us make the Awards column more usable
award <- function(S){
  if(is.na(S)){
    A<-"none"
    return(A)
  }
  S=tolower(S)
  #won and nominated:
  if((grepl("won", S) | grepl("win", S)) && (grepl("nomina", S))){
    A<-"won and nominated"
  }
  #only won:
  else if (grepl("won", S) | grepl("win", S)){
    A<-"won"
  }
  #only nominated:
  else if (grepl("nomina", S)){
    A<-"nomination"
  }
  #no awards or nominations:
  else {(A<-"none")
     return(A)
  }
}
```

Here's our complete formatting function:

```{r 12_02_6_23_2022,message=FALSE}
library(lubridate)
mat1=NULL
format_data <- function(mykey,titles,series){
    data <- get_data_titles_and_series(mykey,titles,series)
    data$Year <- as.numeric(data$Year)
    data$Released <- dmy(data$Released)
    data$Runtime <- as.numeric(gsub(" min","",data$Runtime))
    data$Ratings.Value <- sapply(data$Ratings.Value, FUN=parse_number)
    data$Summary_Awards <- as.factor(sapply(data$Awards, FUN=award))
    data$Metascore <- as.numeric(data$Metascore)
    data$imdbRating <- as.numeric(data$imdbRating)*10
    data$imdbVotes <- as.numeric(gsub(",","",data$imdbVotes))
    data$DVD <- dmy(data$DVD)
    data$BoxOffice <- gsub("\\$","",data$BoxOffice)
    data$BoxOffice <- as.numeric(gsub(",","",data$BoxOffice))
    movie_list<-unique(data$Title)
    
    for (i in movie_list){
      temp=data[is.element(data$Title,i),]
      Ratings.Value_mean<-mean(temp$Ratings.Value)
      Metascore<-unique(temp$Metascore)
      imdbRating<-unique(temp$imdbRating)
      
      if(is.na(Metascore)==TRUE){
        temp$average_rating=(Ratings.Value_mean+imdbRating)/2
      }
      if(is.na(Metascore)==FALSE){
        temp$average_rating=(Ratings.Value_mean+Metascore+imdbRating)/3
      }
      mat1=rbind(mat1,temp)
    }
    return(mat1)
}
```

Here's how I ran it:

```{r 11_40_6_23_2022,eval=FALSE}
formatted_data<-format_data("5c7f9206",titles,series)
str(formatted_data)
```

Here's the tibble I got:

```{r 12_16_6_23_2022,echo=FALSE,warning=FALSE}
formatted_data<-format_data("5c7f9206",titles,series)
str(formatted_data) #this shows you the structure of formatted_data
```

# Contingency Tables

Let's say I wanted to know how many movies are rated "PG-13" in my data set. I would make a contingency table summarizing the ratings for all of the movies: 

```{r 11_45_6_23_2022}
unique_titles<-length(unique(formatted_data$Title))
unique_titles #there are 61 unique movie titles in my data set
A<-data.frame(formatted_data$Title,formatted_data$Rated)
B<-unique(A)
table(B$formatted_data.Rated)
```

As you can see, 9 out of the 61 movies from my data set are "PG-13".

Let's say I wanted to summarize how many movies fell under each genre listed. I would make the contingency table below:

```{r 11_50_6_23_2022}
C<-data.frame(formatted_data$Title,formatted_data$Genre)
D<-unique(C)
table(D$formatted_data.Genre)
```

So, for example, it looks like 7 out of the 61 from my data set are classified under the "Horror, Thriller" genre.

What if we wanted to see both the rating and the genre in the same table? We could make a two-way contingency table:

```{r 12_11_6_24_2022}
table(B$formatted_data.Rated, D$formatted_data.Genre)
```

So, for example, it looks like 5 out of the 61 movies from my data set are rated PG and fall under the "Action, Adventure, Fantasy" genre.

# Bar Plots

Let's say we wanted to find out how many movies each director made in my data set. We could make the bar graph below:

```{r 8_18_6_24_2022,fig.height = 10, fig.width = 10}
C<-data.frame(formatted_data$Title,formatted_data$Director)
D<-unique(C)
g<-ggplot(data = D, aes(x = formatted_data.Director ))
g + geom_bar() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = "Director", title = "Director Count", y="Number of Movies") +
  coord_flip()
```

It looks like George Lucas and Francis Ford Coppola are tied for the most number of movies.

Now, what if we wanted to see, for each director, how many movies either won or was nominated for awards. We could make the bar graph below.

```{r 10_54_6_25_2022, fig.height = 10, fig.width = 10}
C<-data.frame(formatted_data$Title,formatted_data$Director,formatted_data$Summary_Awards)
D<-unique(C)
g<-ggplot(data = D, aes(x = formatted_data.Director ))
g + geom_bar(aes(fill = as.factor(formatted_data.Summary_Awards))) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = "Director", title = "Director Count", y="Number of Movies") +
  scale_fill_discrete(name = NULL) +
  coord_flip()
```

It looks like, for all 4 of George Lucas' movies, all of them won and were nominated for awards.

# Measures of Center and Spread

Let's explore the imdbVotes column. First, let's find the average number of imdbVotes in our entire data set:

```{r 11_38_6_25_2022}
imdbVotes <- unique(formatted_data$imdbVotes)
avg_imdbVotes <- mean(imdbVotes)
avg_imdbVotes
```

On average, every movie in my data set gets 332324 imdbVotes. Now, let's get the average imdbVotes for every genre:

```{r 11_40_6_25_2022}
A<-data.frame(formatted_data$Genre,formatted_data$imdbVotes)
B<-unique(A)
genre_list=unique(B$formatted_data.Genre)
length(genre_list) #there are 30 different genres
```

```{r 12_29_6_25_2022, fig.height = 10, fig.width = 10}
g <- ggplot(B, aes(x = formatted_data.Genre, y = formatted_data.imdbVotes))
g + geom_boxplot(fill = "grey") +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = "Genre", title = "imdbVotes By Genre", y="Number of imdbVotes") +
  coord_flip()



```


```{r 12_28_6_25_2022}
mat1=NULL
avg_imdbVotes_by_genre <- function(){
  for (i in genre_list){
      temp=B[is.element(B$formatted_data.Genre,i),]
      temp$avg_genre<-mean(temp$formatted_data.imdbVotes)
      mat1=rbind(mat1,temp)
  }
  return(mat1)
}
result<-avg_imdbVotes_by_genre()
result<-result %>% select(formatted_data.Genre,avg_genre)
result<-unique(result)
rownames(result)<-1:30
result
```

Now, let's find the 5-number summary for this data set.

```{r 12_14_6_25_2022}
summary(result$avg_genre)
```

So, it looks like the genre with the most number of votes is "Biography, Crime, Drama" and the genre with the least number of votes is "Documentary, Short". Let's make a boxplot so that we can visually see this:

```{r 12_18_6_25_2022}
g <- ggplot(result, aes(y = avg_genre))
g + geom_boxplot(fill = "grey")



```



movie title in mind, like Star Wars. Here's a function you can use to get data from the OMDb API about Star Wars:



This is my .Rmd file. Test. Test 2

```{r plot}
plot(iris)
```